---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r Libraries, include=F}
library(openxlsx)
library(tidyverse)
library(stringr)
library(knitr)
library(ggplot2)
```

```{r LOADING data, include=F, }
# rouen_doc <- read.xlsx("Extraction.xlsx", sheet = 1)
rouen_sej_pmsi <- read.xlsx("Extraction.xlsx", sheet = 2)
rouen_atb <- read.xlsx("Extraction.xlsx", sheet = 3)
rouen_codes_pmsi <- read.xlsx("Extraction.xlsx", sheet = 4)
rouen_codes_ccam <- read.xlsx("Extraction.xlsx", sheet = 5)
rouen_concepts_long <- read.xlsx("Extraction.xlsx", sheet = 6)
rouen_concepts_wide <- read.xlsx("Extraction.xlsx", sheet = 7)
dictionnaire <- read.xlsx("Extraction.xlsx", sheet = 8, rows = 31:52, cols = 1:2)
rouen_nomenclature_ccam <- read.xlsx("Extraction.xlsx", sheet = 9)
```

```{r dm PMSI, include=F}
# Properly formatting to wide format
rouen_codes_pmsi <-
  rouen_codes_pmsi %>%
  group_by(PATID, EVTID) %>%
  summarise(PMSI = str_flatten(PMSI))
# Widening
rouen_codes_pmsi <- rouen_codes_pmsi %>%
  mutate(PMSI = str_split_fixed(PMSI, "\\s", n = 50)) %>%
  apply(2, str_remove, "\\:.+$") %>%
  data.frame()

# Pivoting back to long format to comfortably erase duplicates then back to wide format
rouen_codes_pmsi <-
  pivot_longer(rouen_codes_pmsi, cols = 3:52, values_to = "pmsi") %>%
  arrange(EVTID, pmsi, by_group = T) %>%
  select(-name) %>% # dropping "name" column
  distinct() %>% # dropping duplicates
  group_by(PATID, EVTID) %>%
  summarise(pmsi = str_flatten(pmsi, collapse = ",")) %>%
  mutate(pmsi = str_remove(pmsi, "\\,$|^\\,"))
# Adding PMSI labels
rouen_codes_pmsi$pmsi_label <- case_when(
  str_detect(rouen_codes_pmsi$pmsi, "K047") ~ "Abces periapical",
  str_detect(rouen_codes_pmsi$pmsi, "K046") ~ "Abces periapical avec fistule",
  str_detect(rouen_codes_pmsi$pmsi, "K052") ~ "Periodontite aigue",
  str_detect(rouen_codes_pmsi$pmsi, "K053") ~ "Periodondite",
  str_detect(rouen_codes_pmsi$pmsi, "K02.") ~ "Carie",
  str_detect(rouen_codes_pmsi$pmsi, "K088") ~ "Autres affections precisees des dents et du parodonte",
  str_detect(rouen_codes_pmsi$pmsi, "K089") ~ "Affection des dents et du parodonte, sans precision",
  str_detect(rouen_codes_pmsi$pmsi, "K040|K045|K048") ~ "Maladies de la pulpe et des tissu périapicaux, autres et sans précision",
  str_detect(rouen_codes_pmsi$pmsi, "K050|K056") ~ "Maladie périodontale, sans précision",
  str_detect(rouen_codes_pmsi$pmsi, "K068") ~ "Affection de la gencive et de la crête alvéolaire édentée, sans précision"
)
# names to lower case
names(rouen_codes_pmsi) <- tolower(names(rouen_codes_pmsi))

# Now that DPs always precede DS, we can remove labels to make codes more readable
rouen_codes_pmsi$pmsi <- rouen_codes_pmsi$pmsi %>% str_remove_all("DS\\'|DP\\'|DD\\'")
```

```{r dm CCAM, include = FALSE}
rouen_codes_ccam <-
  left_join(rouen_codes_ccam, rouen_nomenclature_ccam[c(1, 2)], by = c("CODEACTE" = "Code"))
# names to lower case
names(rouen_codes_ccam) <- tolower(names(rouen_codes_ccam))
```

```{r dm SEJOURS, include = F}
# names to lower case
names(rouen_sej_pmsi) <- tolower(names(rouen_sej_pmsi))

# Preparing dictionary for join
names(dictionnaire) <- c("code", "label")
dictionnaire$code <- as.character(dictionnaire$code)

# Joining labels to main table
rouen <-
  left_join(rouen_sej_pmsi, dictionnaire, by = c("modeent" = "code")) %>%
  rename(modeent_label = label) %>%
  left_join(dictionnaire, by = c("modesort" = "code")) %>%
  rename(modesort_label = label)

# Replacing "retour du domicile" with "retour AU domicile")
rouen$modesort_label[which(rouen$modesort_label == "Du domicile")] <- "Au domicile"
```

```{r dm joining tables, include = F}
names(rouen_codes_pmsi) <- tolower(names(rouen_codes_pmsi))
# Adding pmsi codes table
rouen <- left_join(rouen, rouen_codes_pmsi)
# Adding ccam acts table
rouen <-
  left_join(rouen, rouen_codes_ccam[c("patid", "evtid", "codeacte", "texte")]) %>%
  rename(acte = codeacte, acte_label = texte)

rouen <- distinct(rouen) # erasing eventual duplicates
```

# Descriptive analysis

```{r include=F}
rouen$sejdur <- as.numeric(rouen$sejdur)

unique_ids <- length(unique(rouen$patid))
unique_stays <- length(unique(rouen$evtid))

# How many stays started with an arrival from personal home
entry_mode <- rouen %>%
  select(evtid, sejum, modeent_label) %>%
  filter(sejum == "URGE" & modeent_label == "Du domicile") %>%
  distinct() %>%
  nrow()
```

We have `r unique_ids` **unique patients** and `r unique_stays` **unique hospital stays/emergency ward visits** so basically every patient consulted only once for dental problems, at the hospital, over the period of interest. 
`r round((entry_mode/unique_stays)*100,2)`% of the consultations/stays started with a visit to the emergency ward, coming from domicile as it was a selection criteria. 

Patient consulted for:
```{r DP de consultation, echo=F}
rouen %>%
  select(evtid, pmsi_label) %>%
  distinct() %>%
  count(pmsi_label) %>%
  mutate(freq = round(prop.table(n) * 100, 2)) %>%
  arrange(desc(n)) %>%
  kable(col.names = c("Main diagnosis", "n", "freq"))

rouen %>%
  select(evtid, pmsi_label) %>%
  distinct() %>%
  count(pmsi_label) %>%
  ggplot(aes(y = factor(pmsi_label, levels = pmsi_label[order(n)]), x = n)) +
  geom_bar(stat = "identity")+
    labs(y="Main diagnosis")
    
    
    
```

**Mean lenght of stay** was `r round(mean(rouen$sejdur,na.rm=T),2)` days.
