---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r Libraries, include=F}
library(openxlsx)
library(tidyverse)
library(stringr)
library(knitr)
library(ggplot2)
```
```{r LOADING data, include=F, }
# rouen_doc <- read.xlsx("Extraction.xlsx", sheet = 1)
rouen_sej_pmsi <- read.xlsx("Extraction.xlsx", sheet = 2)
rouen_atb <- read.xlsx("Extraction.xlsx", sheet = 3)
rouen_codes_pmsi <- read.xlsx("Extraction.xlsx", sheet = 4)
rouen_codes_ccam <- read.xlsx("Extraction.xlsx", sheet = 5)
rouen_concepts_long <- read.xlsx("Extraction.xlsx", sheet = 6)
rouen_concepts_wide <- read.xlsx("Extraction.xlsx", sheet = 7)
dictionnaire <- read.xlsx("Extraction.xlsx", sheet = 8, rows = 31:52, cols = 1:2)
rouen_nomenclature_ccam <- read.xlsx("Extraction.xlsx", sheet = 9)
```
```{r dm PMSI, include=F}
# Properly formatting to wide format
rouen_codes_pmsi <-
  rouen_codes_pmsi %>%
  group_by(PATID, EVTID) %>%
  summarise(PMSI = str_flatten(PMSI))
# Widening
rouen_codes_pmsi <- rouen_codes_pmsi %>%
  mutate(PMSI = str_split_fixed(PMSI, "\\s", n = 50)) %>%
  apply(2, str_remove, "\\:.+$") %>%
  data.frame()

# Pivoting back to long format to comfortably erase duplicates then back to wide format
rouen_codes_pmsi <-
  pivot_longer(rouen_codes_pmsi, cols = 3:52, values_to = "pmsi") %>%
  arrange(EVTID, pmsi, by_group = T) %>%
  select(-name) %>% # dropping "name" column
  distinct() %>% # dropping duplicates
  group_by(PATID, EVTID) %>%
  summarise(pmsi = str_flatten(pmsi, collapse = ",")) %>%
  mutate(pmsi = str_remove(pmsi, "\\,$|^\\,"))
# Adding PMSI labels
rouen_codes_pmsi$pmsi_label <- case_when(
  str_detect(rouen_codes_pmsi$pmsi, "K047") ~ "Abces periapical",
  str_detect(rouen_codes_pmsi$pmsi, "K046") ~ "Abces periapical avec fistule",
  str_detect(rouen_codes_pmsi$pmsi, "K052") ~ "Periodontite aigue",
  str_detect(rouen_codes_pmsi$pmsi, "K053") ~ "Periodondite",
  str_detect(rouen_codes_pmsi$pmsi, "K02.") ~ "Carie",
  str_detect(rouen_codes_pmsi$pmsi, "K088") ~ "Autres affections precisees des dents et du parodonte",
  str_detect(rouen_codes_pmsi$pmsi, "K089") ~ "Affection des dents et du parodonte, sans precision",
  str_detect(rouen_codes_pmsi$pmsi, "K040|K045|K048") ~ "Maladies de la pulpe et des tissu périapicaux, autres et sans précision",
  str_detect(rouen_codes_pmsi$pmsi, "K050|K056") ~ "Maladie périodontale, sans précision",
  str_detect(rouen_codes_pmsi$pmsi, "K068") ~ "Affection de la gencive et de la crête alvéolaire édentée, sans précision"
)
# names to lower case
names(rouen_codes_pmsi) <- tolower(names(rouen_codes_pmsi))

# Now that DPs always precede DS, we can remove labels to make codes more readable
rouen_codes_pmsi$pmsi <- rouen_codes_pmsi$pmsi %>% str_remove_all("DS\\'|DP\\'|DD\\'")
```
```{r dm CCAM, include = FALSE}
rouen_codes_ccam <-
  left_join(rouen_codes_ccam, rouen_nomenclature_ccam[c(1, 2)], by = c("CODEACTE" = "Code"))
# names to lower case
names(rouen_codes_ccam) <- tolower(names(rouen_codes_ccam))
```
```{r dm SEJOURS, include = F}
# names to lower case
names(rouen_sej_pmsi) <- tolower(names(rouen_sej_pmsi))

# Preparing dictionary for join
names(dictionnaire) <- c("code", "label")
dictionnaire$code <- as.character(dictionnaire$code)

# Joining labels to main table
rouen <-
  left_join(rouen_sej_pmsi, dictionnaire, by = c("modeent" = "code")) %>%
  rename(modeent_label = label) %>%
  left_join(dictionnaire, by = c("modesort" = "code")) %>%
  rename(modesort_label = label)

# Replacing "retour du domicile" with "retour AU domicile")
rouen$modesort_label[which(rouen$modesort_label == "Du domicile")] <- "Au domicile"
```
```{r dm joining tables, include = F}
names(rouen_codes_pmsi) <- tolower(names(rouen_codes_pmsi))
# Adding pmsi codes table
rouen <- left_join(rouen, rouen_codes_pmsi)
# Adding ccam acts table
rouen <-
  left_join(rouen, rouen_codes_ccam[c("patid", "evtid", "codeacte", "texte")]) %>%
  rename(acte = codeacte, acte_label = texte)

rouen <- distinct(rouen) # erasing eventual duplicates
```
# Descriptive analysis
```{r include=F}
rouen$sejdur <- as.numeric(rouen$sejdur)

unique_ids <- length(unique(rouen$patid))
unique_stays <- length(unique(rouen$evtid))

rouen$patage <- as.numeric(rouen$patage)
mean_age <- round(mean(rouen$patage, na.rm = T), 2)
female_mean_age <- round(y(rouen$patage, rouen$patsex, mean, na.rm = T)[["F"]], 2)
male_mean_age <- round(by(rouen$patage, rouen$patsex, mean, na.rm = T)[["M"]], 2)

# How many stays started with an arrival from personal home
entry_mode <- rouen %>%
  select(evtid, sejum, modeent_label) %>%
  filter(sejum == "URGE" & modeent_label == "Du domicile") %>%
  distinct() %>%
  nrow()
```
We have `r unique_ids` **unique patients** and `r unique_stays` **unique hospital stays/emergency ward visits** so basically every patient consulted only once for dental problems, at the hospital, over the period of interest which is the calendar year 2021.

## Age
**Mean age**  for both sex combined is `r mean_age` years (*sd* = `r round(sd(rouen$patage),2)`) while
it's `r female_mean_age` (*sd* = `r round(sd(rouen$patage[rouen$patsex=="F"]),2)`) for women and `r male_mean_age` (*sd* = `r round(sd(rouen$patage[rouen$patsex=="M"]),2)`)for men.
```{r age distribution, echo = F}
rouen %>%
  select(patid, evtid, patage, patsex) %>%
  distinct() %>%
  ggplot() +
  geom_histogram(aes(x = patage), binwidth = 5, color = "white", boundary = .5) +
  scale_x_continuous(breaks = seq(0, 100, 5)) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  labs(
    title = "Age distribution",
    x = "Age",
    y = "Frequency"
  )

rouen %>%
  select(patid, evtid, patage, patsex) %>%
  distinct() %>%
  ggplot(aes(x = patage)) +
  geom_histogram(binwidth = 5, aes(fill = patsex), alpha = .8, position = "identity", boundary = .5) +
  # geom_density(aes(y = ..density.. * (nrow(rouen) * 5), color = patsex), size = 1)+
  scale_x_continuous(breaks = seq(0, 100, 5)) +
  scale_y_continuous(breaks = seq(0, 100, 5)) +
  labs(
    title = "Age distribution by sex",
    x = "Age",
    y = "Frequency",
    fill = "Sex"
  )
```

## Consultations
`r round((entry_mode/unique_stays)*100,2)`% of the consultations/stays started with a visit to the emergency ward, coming from domicile as it was a selection criteria. 

Patient consulted for:
```{r DP de consultation, echo=F}
rouen %>%
  select(evtid, pmsi_label) %>%
  distinct() %>%
  count(pmsi_label) %>%
  mutate(freq = round(prop.table(n) * 100, 2)) %>%
  arrange(desc(n)) %>%
  kable(col.names = c("Main diagnosis", "n", "freq"))

rouen %>%
  select(evtid, pmsi_label) %>%
  distinct() %>%
  count(pmsi_label) %>%
  ggplot(aes(y = factor(pmsi_label, levels = pmsi_label[order(n)]), x = n, 
             label = c("K047", "K046", "K068", "K089", "K088", "K02*", "K056", 
                       "K049", "K053", "K052"))) +
  geom_bar(stat = "identity") +
    geom_label(hjust="left", nudge_x = 5)+
  labs(y = "Main diagnosis")
```

## Stays
**Mean lenght of stay** was `r round(mean(rouen$sejdur,na.rm=T),2)` days.
```{r echo=F}
rouen$pmsi <- str_replace_all(rouen$pmsi, "K040|K045|K048", "K049")
rouen$pmsi <- str_replace_all(rouen$pmsi, "K050|K056", "K056")

cbind(
  "Code" = c("K047", "K046", "K052", "K053", "K02*", "K088", "K089", "K049", "K056", "K068"),
  "Label" = c("Abces periapical", "Abces periapical avec fistule", "Periodontite aigue",
    "Periodondite", "Carie", "Autres affections precisees des dents et du parodonte",
    "Affection des dents et du parodonte, sans precision",
    "Maladies de la pulpe et des tissu périapicaux, autres et sans précision",
    "Maladie périodontale, sans précision",
    "Affection de la gencive et de la crête alvéolaire édentée, sans précision"
  )
) %>% data.frame %>% arrange(Code) %>% kable()
```
